Frequentist Statistic

### 2016-12-28 ab_tests/frequentist_ab_test.Rmd
rerun rmarkdown, add license

---
title: "Frequentist A/B test"
author: "Ming-Yu Liu"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: pygments
---

@@

PlotPower <- function( size, min_diff ){

	size_a <- size_b <- size * 0.5 # size are assumed to be equal
	p_a <- 0.08 # baseline
	p_b <- p_a + min_diff
	count_a  <- size_a * p_a
	count_b  <- size_b * p_b
	p_pooled <- ( count_a  + count_b ) / ( size_a + size_b )
	Z <- ( p_b - p_a ) / sqrt( p_pooled * ( 1 - p_pooled ) * ( 1 / size_a + 1 / size_b ) )

@@

	x <- seq( -4, 6, 0.1 ) # use for generating the x axis of the normal distribution
	data <- data.frame( x = x, y1 = dnorm( x, mean1, 1 ), y2 = dnorm( x, mean2, 1 ) )

	plot <- ggplot( data, aes( x = x ) ) +
			geom_line( aes( y = y1, colour = 'H0 is true' ), size = 1.2 ) +
			geom_line( aes( y = y2, colour = 'H1 is true' ), size = 1.2 ) +
			geom_area( aes( y = y1, x = ifelse( x > 1.65, x, NA ) ), fill = 'black' ) +
			geom_area( aes( y = y2, x = ifelse( x > 1.65, x, NA ) ), fill = 'blue', alpha = 0.3 ) +
			labs( x = '', y = '', title = sprintf( 'p1 = %s, p2 = %s, size = %d', p_a, p_b, size ) ) + 
			theme( legend.title = element_blank() ) +
			scale_colour_manual( breaks = c( "H0 is true", "H1 is true" ), 
								 values = c( "blue", "red" ) )
	return(plot)
}
```

@@

PlotPower( size = 5000, min_diff = 0.02 )

@@

# smaller N
PlotPower( size = 2500, min_diff = 0.02 )

# smaller probability difference
PlotPower( size = 5000, min_diff = 0.01 )

@@

# calculate the the required sample size for reaching the 
# range detectable difference (dd)
dd <- seq( from = 0.01, to = 0.03, by = 0.0001 )
result <- matrix( nrow = length(dd), ncol = 2 )
result[ , 1 ] <- dd
for( i in 1:length(dd) ){
	result[ i, 2 ] <- power.prop.test( p1 = baseline, p2 = baseline - dd[i], 
									   power = power, sig.level = sig_level,
									   alternative = "two.sided" )$n
}

result <- data.table(result)
setnames( result, c( 'dd', 'n' ) )

ggplot(data = result, aes( x = dd, y = n ) ) +
geom_line() + ylab("required sample size") + xlab("Detectable difference") + 
scale_x_continuous( labels = comma )

@@

result <- prop.test( c( count_control, count_experiment ), 
					 c( sizes_control, sizes_experiment ) )
result

# do the computation ourselves to see that the confidence interval matches
# compute the probability of each group and the standard error
p1 <- count_control / sizes_control
p2 <- count_experiment / sizes_experiment
se <- sqrt( p1 * ( 1 - p1 ) / sizes_control + p2 * ( 1 - p2 ) / sizes_experiment )

# 95 percent confidence interval's z score
conf_level <- 0.95
zscore <- qnorm( conf_level + ( 1 - conf_level ) / 2 )
conf_int <- abs(p2 - p1) + c( -1, 1 ) * zscore * se
conf_int

@@

ggplot( examples, aes( mean, scenario, color = scenario ) ) + 
geom_point() + 
geom_errorbarh( aes( xmin = lower, xmax = upper ), height = 0.1 ) + 
geom_vline( xintercept = 0, color = "black" ) + 
geom_vline( xintercept = delta, color = "blue", linetype = "dotted" ) + 
geom_vline( xintercept = -delta, color = "blue", linetype = "dotted" ) +
scale_color_discrete( breaks = as.character(1:6) ) +  
labs( title = "Different Scenarios of Confidence Interval",
	  x = "confidence interval" )

@@

group1 <- 64454
group2 <- 61818
SanityCheck <- function( group1, group2 )
{
	# 95 percent confidence interval = qnorm(0.95 + (1-0.95) / 2 )
	n <- group1 + group2
	confidence <- n * 0.5 + c( -1, 1 ) * qnorm(0.975) * sqrt( n * 0.5 * 0.5 ) 
	return(confidence)
}
( sanity <- SanityCheck( group1, group2 ) )

@@

@@

- [Stitchfix's Blog: So, You Need a Statistically Significant Sample?](http://multithreaded.stitchfix.com/blog/2015/05/26/significant-sample/)
- [Z-score distribution table](http://www.utdallas.edu/dept/abp/zscoretable.pdf)
- [Minitab's Blog: How to Correctly Interpret P Values](http://blog.minitab.com/blog/adventures-in-statistics/how-to-correctly-interpret-p-values)

sessionInfo()

>>>>

---
title: "Frequentist A/B test"
author: "Ethen Liu"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: pygments
---

@@

plot_power <- function(size, min_diff) {

	size_a <- size_b <- size * 0.5 # size are assumed to be equal
	p_a <- 0.08 # baseline
	p_b <- p_a + min_diff
	count_a  <- size_a * p_a
	count_b  <- size_b * p_b
	p_pooled <- (count_a  + count_b) / (size_a + size_b)
	Z <- (p_b - p_a) / sqrt( p_pooled * (1 - p_pooled) * (1 / size_a + 1 / size_b) )

@@

	x <- seq(-4, 6, 0.1) # use for generating the x axis of the normal distribution
	data <- data.frame( x = x, y1 = dnorm(x, mean1, 1), y2 = dnorm(x, mean2, 1) )

	plot <- ggplot( data, aes(x = x) ) +
			geom_line( aes( y = y1, colour = 'H0 is true' ), size = 1.2 ) +
			geom_line( aes( y = y2, colour = 'H1 is true' ), size = 1.2 ) +
			geom_area( aes( y = y1, x = ifelse(x > 1.65, x, NA) ), fill = 'black' ) +
			geom_area( aes( y = y2, x = ifelse(x > 1.65, x, NA) ), fill = 'blue', alpha = 0.3 ) +
			labs( x = '', y = '', title = sprintf('p1 = %s, p2 = %s, size = %d', p_a, p_b, size) ) + 
			theme( legend.title = element_blank() ) +
			scale_colour_manual( breaks = c("H0 is true", "H1 is true"), 
								 values = c("blue", "red") )
	return(plot)
}
```

@@

plot_power(size = 5000, min_diff = 0.02)

@@

# smaller N
plot_power(size = 2500, min_diff = 0.02)

# smaller probability difference
plot_power(size = 5000, min_diff = 0.01)

@@

# calculate the the required sample size for reaching the 
# range detectable difference (dd)
dd <- seq(from = 0.01, to = 0.03, by = 0.0001)
result <- matrix(nrow = length(dd), ncol = 2)
result[, 1] <- dd
for( i in 1:length(dd) ) {
	result[i, 2] <- power.prop.test( p1 = baseline, p2 = baseline - dd[i], 
									   power = power, sig.level = sig_level,
									   alternative = "two.sided" )$n
}

result <- data.table(result)
setnames( result, c('dd', 'n') )

ggplot(data = result, aes(x = dd, y = n) ) +
geom_line() + ylab("required sample size") + xlab("Detectable difference") + 
scale_x_continuous(labels = comma)

@@

result <- prop.test( c(count_control, count_experiment), 
					 c(sizes_control, sizes_experiment) )
result

# do the computation ourselves to see that the confidence interval matches
# compute the probability of each group and the standard error
p1 <- count_control / sizes_control
p2 <- count_experiment / sizes_experiment
se <- sqrt( p1 * (1 - p1) / sizes_control + p2 * (1 - p2) / sizes_experiment )

# 95 percent confidence interval's z score
conf_level <- 0.95
zscore <- qnorm( conf_level + (1 - conf_level) / 2 )
conf_int <- abs(p2 - p1) + c(-1, 1) * zscore * se
conf_int

@@

ggplot( examples, aes(mean, scenario, color = scenario) ) + 
geom_point() + 
geom_errorbarh( aes(xmin = lower, xmax = upper), height = 0.1 ) + 
geom_vline(xintercept = 0, color = "black") + 
geom_vline(xintercept = delta, color = "blue", linetype = "dotted") + 
geom_vline(xintercept = -delta, color = "blue", linetype = "dotted") +
scale_color_discrete( breaks = as.character(1:6) ) +  
labs( title = "Different Scenarios of Confidence Interval",
	  x = "confidence interval" )

@@

group1 <- 64454
group2 <- 61818
sanity_check <- function(group1, group2) {
	# 95 percent confidence interval = qnorm(0.95 + (1-0.95) / 2 )
	n <- group1 + group2
	confidence <- n * 0.5 + c(-1, 1) * qnorm(0.975) * sqrt(n * 0.5 * 0.5) 
	return(confidence)
}
( sanity <- sanity_check(group1, group2) )

@@

```{r}
devtools::session_info()
```

@@

- [Blog: So, You Need a Statistically Significant Sample?](http://multithreaded.stitchfix.com/blog/2015/05/26/significant-sample/)
- [Blog: How to Correctly Interpret P Values](http://blog.minitab.com/blog/adventures-in-statistics/how-to-correctly-interpret-p-values)
- [Z-score distribution table](http://www.utdallas.edu/dept/abp/zscoretable.pdf)

### 2016-12-28 articles/articles.Rmd
rerun rmarkdown, add license

- What factors can increase your data science salary?. [[Springboard](https://www.springboard.com/blog/highest-data-scientist-salary-possible/?utm_campaign=Data%2BElixir&utm_medium=email&utm_source=Data_Elixir_90)]
- 5 Qualities You Need to Show in the Job Search. [[glassdoor](https://www.glassdoor.com/blog/qualities-show-off/?utm_source=newsletter&utm_medium=email&utm_content=Top5Jobs_US16&utm_campaign=May2016_US)]
    - Leadership ability. e.g. As a result of an incentive program I spearheaded, our sales team surpassed its quarterly sales goal by 15 percent.
    - Flexibility: keep track of times when you’ve had to assume multiple roles or responsibilities outside of your own.
    - Coachability: Be open and honest about where you fall short and how you’re actively improving on those shortcomings.
- Labels are for cans not for people. [[Facebook Videos](https://www.facebook.com/morokcom/videos/237500609921393/)]
- Land any job you want. [[blog](http://landanyjobyouwant.com/)]
- The career advice no one tells you. [[Quartz](http://qz.com/640112/the-career-advice-no-one-tells-you/)]
- Wanted in College Graduates: Tolerance for Ambiguity. [[Linkedin Pulse](https://www.linkedin.com/pulse/wanted-college-graduates-tolerance-ambiguity-jeff-selingo)]

@@

>>>>

- 2016.12.27 | Crushed it! Landing a data science job. [[Blog](http://www.erinshellman.com/crushed-it-landing-a-data-science-job/)]
- 2016.12.27 | http://treycausey.com/data_science_interviews.html. [[Blog](http://treycausey.com/data_science_interviews.html)]
- What factors can increase your data science salary?. [[Springboard](https://www.springboard.com/blog/highest-data-scientist-salary-possible/?utm_campaign=Data%2BElixir&utm_medium=email&utm_source=Data_Elixir_90)]
- 5 Qualities You Need to Show in the Job Search. [[glassdoor](https://www.glassdoor.com/blog/qualities-show-off/?utm_source=newsletter&utm_medium=email&utm_content=Top5Jobs_US16&utm_campaign=May2016_US)]
    - Leadership ability. e.g. As a result of an incentive program I spearheaded, our sales team surpassed its quarterly sales goal by 15 percent.
    - Flexibility: keep track of times when you’ve had to assume multiple roles or responsibilities outside of your own.
    - Coachability: Be open and honest about where you fall short and how you’re actively improving on those shortcomings.
- Labels are for cans not for people. [[Facebook Videos](https://www.facebook.com/morokcom/videos/237500609921393/)]
- Land any job you want. [[blog](http://landanyjobyouwant.com/)]
- The career advice no one tells you. [[Quartz](http://qz.com/640112/the-career-advice-no-one-tells-you/)]
- Wanted in College Graduates: Tolerance for Ambiguity. [[Linkedin Pulse](https://www.linkedin.com/pulse/wanted-college-graduates-tolerance-ambiguity-jeff-selingo)]

@@

- 2016.12.27 | Sharing Your Side Projects Online and Making Your Github the Best Résumé. [[Youtube](https://www.youtube.com/watch?v=uRul8QdYvqQ)]
    - Include a license to the project. there's even a [website](http://choosealicense.com/) that walks you through the copying and pasting process
    - [Binder](http://mybinder.org/) is a project that allows you to open up the jupyter notebook in a virual environment (specifying all the requirements in a requirements.txt file) and run it online without needing to download it
- 2016.12.24 | WHAT I LEARNED RECREATING ONE CHART USING 24 TOOLS. [[Blog](https://source.opennews.org/en-US/articles/what-i-learned-recreating-one-chart-using-24-tools/?utm_campaign=Data%2BElixir&utm_medium=web&utm_source=Data_Elixir_110)]

### 2016-12-28 frequentist_statistics/frequentist_statistics.Rmd
fix typo
rerun rmarkdown, add license

---
title: "Frequentist Statistic"
author: "Ming-Yu Liu"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: pygments
---

@@

library(knitr)
opts_chunk$set( message = FALSE, warning = FALSE )

@@

$$\text{point estimate} \pm z * SE $$

@@

# e.g. the positive and negative critical value for 95%
cl <- 0.95
c( -1, 1 ) * qnorm( ( 1 - cl ) / 2 )

@@

3.2 + c( 1, -1 ) * qnorm( ( 1 - 0.95 ) / 2 ) * 1.74 / sqrt(50)

@@

# z-score is computed as ( value - mean ) / standard error
upper <- ( 42 - 40 ) / se
lower <- ( 38 - 40 ) / se
pnorm(upper) - pnorm(lower)

@@

# 1.65 is the critical value, z, for 90%
( 1.65 * 18 / 4 ) ^ 2

@@

# 1.65 is the critical value, z, for 90%
( 1.65 * 18 / 2 ) ^ 2

@@

# We're able to compute the confidence interval using the following formula
3.2 + c( 1, -1 ) * qnorm( ( 1 - 0.95 ) / 2 ) * 1.74 / sqrt(50)

@@

# z score
z <- ( 3.2 - 3 ) / se
pvalue <- pnorm( z, lower.tail = FALSE )
pvalue

@@

If the p-value is low, and by low we mean lower than the significance level $\alpha$, which we usually set at 0.05, but we can certainly change it as well, we say that it would be very unlikely to observe the data if the null hypothesis were true, and therefore we reject the null hypothesis. If on the other hand the p-value is higher than $\alpha$, we say that it is indeed likely to observe the data even if the null hypothesis were true. And hence, we would not reject the null hypothesis.

So back to our original question. Our null hypothesis was that college students on average have three exclusive relationships, versus, the alternative was that that number was something greater than 3.

In our case there is a 20.9% chance that a random sample of 50 college students would still yield a sample mean of 3.2 or higher. Since this is a pretty high probability, we think that a sample mean of 3.2 or more exclusive relationships is likely to happen simply by chance.

@@

@@

df <- min( n1 - 1, n2 - 1 )
critical <- qt( ( 1 - 0.95 ) / 2, df )
se <- sqrt( s1 ^ 2 / n2 + s2 ^ 2 / n2 )

# confidence interval
print( ( x1 - x2 ) + c( 1, -1 ) * critical * se )

# hypothesis test
tscore <- ( x1 - x2 ) / se
pt( tscore, df, lower.tail = FALSE ) * 2

@@

@@

# built-in data PlantGrowth
plant_df <- PlantGrowth
plant_df$group <- factor( plant_df$group,
						  labels = c( 'Control', 'Treatment 1', 'Treatment 2' ) )

ggplot( plant_df, aes( x = group, y = weight ) ) + 
geom_boxplot()

@@

anova_test <- aov( weight ~ group, data = plant_df )
summary(anova_test)

@@

n  <- 200
p  <- 0.9
se <- sqrt( p * ( 1 - p ) / n )

zscore <- ( 0.95 - p ) / se
pnorm( zscore, lower.tail = FALSE )

@@

# here, p and 1 - p is the same, thus we can use p squared
( qnorm( 0.95 + ( 1 - ci ) / 2 ) * p / me ) ^ 2

@@

n <- 1983
p <- 0.5
se <- sqrt(  p * ( 1 - p ) / n )
zscore <- ( 0.6 - p ) / se
pnorm( zscore, lower.tail = FALSE )

@@

obs <- c( 1920, 347, 19, 84, 130 )

# expected probability distribution
prob <- c( 0.8029, 0.1206, 0.0079, 0.0292, 0.0394 )
chisq.test( x = obs, p = prob )

@@

# consider different sample size
# first a sample size of 100
z1 <- ( x - mu ) / ( sd / sqrt(100) )
print(z1)

# then inflate it to 10000 and 
# see the difference between the z score,
# recall that the higher the z score the lower
# our p-value
z2 <- ( x - mu ) / ( sd / sqrt(10000) )
print(z2)

@@

power.t.test( power = 0.8, sig.level = 0.05, delta = 3, sd = 12 )

@@

@@

- [The Theorem Every Data Scientist Should Know](http://www.jeannicholashould.com/the-theorem-every-data-scientist-should-know-2.html)
- [RBlogger: One-way Analysis of Variance](http://www.r-bloggers.com/one-way-analysis-of-variance-anova/)
- [Power and sample size calculation for t test in R](http://users.stat.umn.edu/~corbett/classes/3022/Power_and%20_sample_size_for_t_in_R.txt)
- [Coursera: Inference Statistics](https://www.coursera.org/learn/inferential-statistics-intro)

>>>>

---
title: "Frequentist Statistic"
author: "Ethen Liu"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: pygments
---

@@

library(knitr)
opts_chunk$set(message = FALSE, warning = FALSE)

@@

$$\text{point estimate} \pm z * SE$$

@@

# e.g. the positive and negative critical value for 95%
cl <- 0.95
c(-1, 1) * qnorm( (1 - cl) / 2 )

@@

3.2 + c(1, -1) * qnorm( (1 - cl) / 2 ) * 1.74 / sqrt(50)

@@

# z-score is computed as (value - mean) / standard error
upper <- (42 - 40) / se
lower <- (38 - 40) / se
pnorm(upper) - pnorm(lower)

@@

# 1.65 is the critical value, z, for 90%
(1.65 * 18 / 4) ^ 2

@@

# 1.65 is the critical value, z, for 90%
(1.65 * 18 / 2) ^ 2

@@

# We're able to compute the confidence interval using the following formula
3.2 + c(1, -1) * qnorm( (1 - 0.95) / 2 ) * 1.74 / sqrt(50)

@@

# z score
z <- (3.2 - 3) / se
pvalue <- pnorm(z, lower.tail = FALSE)
pvalue

@@

If the p-value is low, and by low we mean lower than the significance level $\alpha$, which we usually set at 0.05, but we can certainly change it as well, we say that it would be very unlikely to observe the data if the null hypothesis were true, and therefore we reject the null hypothesis. If on the other hand the p-value is higher than $\alpha$, we say that it is indeed likely to observe the data if the null hypothesis were true. And hence, we would not reject the null hypothesis.

Back to our original question. Our null hypothesis was that college students on average have 3 exclusive relationships, and the alternative was that that number was something greater than 3. From the result we saw that there was a 20.9% chance that a random sample of 50 college students would still yield a sample mean of 3.2 or higher. Since this is a pretty high probability, we think that a sample mean of 3.2 or more exclusive relationships is likely to happen simply by chance.

@@

p-value is an extremely important concept in frequentist statistics. For those interested, this blog post so covers it with an pizza store example. [Blog: What Are P-Values?](https://prateekvjoshi.com/2013/12/07/what-are-p-values/)

@@

df <- min(n1 - 1, n2 - 1)
critical <- qt( (1 - 0.95) / 2, df )
se <- sqrt(s1 ^ 2 / n2 + s2 ^ 2 / n2)

# confidence interval
print( (x1 - x2) + c(1, -1) * critical * se )

# hypothesis test
tscore <- (x1 - x2) / se
pt(tscore, df, lower.tail = FALSE) * 2

@@

```{r, message=FALSE, warning=FALSE}

@@

# built-in data PlantGrowth
plant_df <- PlantGrowth
plant_df$group <- factor( plant_df$group,
						  labels = c('Control', 'Treatment 1', 'Treatment 2') )

ggplot( plant_df, aes(x = group, y = weight) ) + 
geom_boxplot()

@@

anova_test <- aov(weight ~ group, data = plant_df)
summary(anova_test)

@@

n  <- 200
p  <- 0.9
se <- sqrt( p * (1 - p) / n )

zscore <- (0.95 - p) / se
pnorm(zscore, lower.tail = FALSE)

@@

# here, p and 1 - p is the same, thus we can use p squared
( qnorm( 0.95 + (1 - ci) / 2 ) * p / me ) ^ 2

@@

n <- 1983
p <- 0.5
se <- sqrt(  p * (1 - p) / n )
zscore <- (0.6 - p) / se
pnorm(zscore, lower.tail = FALSE)

@@

obs <- c(1920, 347, 19, 84, 130)

# expected probability distribution
prob <- c(0.8029, 0.1206, 0.0079, 0.0292, 0.0394)
chisq.test(x = obs, p = prob)

@@

# consider different sample size
# first a sample size of 100
z1 <- (x - mu) / ( sd / sqrt(100) )
print(z1)

# then inflate it to 10000 and 
# see the difference between the z score,
# recall that the higher the z score the lower
# our p-value
z2 <- (x - mu) / ( sd / sqrt(10000) )
print(z2)

@@

power.t.test(power = 0.8, sig.level = 0.05, delta = 3, sd = 12)

@@

# R Session Information

```{r}
devtools::session_info()
```

@@

- [Coursera: Inference Statistics](https://www.coursera.org/learn/inferential-statistics-intro)
- [Blog: What Are P-Values?](https://prateekvjoshi.com/2013/12/07/what-are-p-values/)
- [Blog: One-way Analysis of Variance](http://www.r-bloggers.com/one-way-analysis-of-variance-anova/)
- [Blog: The Theorem Every Data Scientist Should Know](http://www.jeannicholashould.com/the-theorem-every-data-scientist-should-know-2.html)
- [Notes: Power and sample size calculation for t test in R](http://users.stat.umn.edu/~corbett/classes/3022/Power_and%20_sample_size_for_t_in_R.txt)
